---
title: "Initial Korean covid 19"
author: "Genevieve Berumen"
date: "10/24/23"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

# Method
According to the data found under data, we are trying to determine if there is a difference in the category Sex (male/female) about COVID-19. We could prove if the virus is more harmful to men or women based on the percentage of people from each group that passes away after contracting it. We can reach this conclusion using Cat~Cat helper files from the CandP zip to help us decipher these statistics.



## Graphical results - use ggplot commands to do this  
```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
  
# the group_by followed by summarize(count=n())
basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot
basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
  
When viewing the first chart, it shows that more women got COVID than men. When looking at the second graph, the results show that more men were being killed than women in the whole group of people that got COVID.     


## Numerical results
```{r}
table2 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table2)
colPerc(table2)
```
These results confirmed the data we saw in the ggplot results. Which was that women got COVID at a higher rate than men. However, of the people who got COVID, more men died than women.

## Inferential results 
```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

Based on the Chisquare and Fisher test, the result shows that if you are a man, you are twice as likely to die from COVID as a woman.   	

## Conclusions

